A Hitchhiker's Guide to Reproducible Research in R

### 2021-07-13 manuscript.Rmd

This tutorial is aimed at researchers who intend to ensure the reproducibility of a research project and are willing to make relevant code, data, and materials available, whether publicly or on request.
We pursue two aims.
First, the present paper introduces the reader to a workflow that ensures reproducibility by building on several technical solutions originating in the software engineering community that eliminate four common threats to reproducibility [see @Peikert2019 for details].
Specifically, we demonstrate how the R package `repro` supports researchers in implementing these solutions.
The second objective is to discuss opportunities of such workflows to improve not only how research outputs are created and recreated but also how research itself is conducted.
One such opportunity is an alternative preregistration <!--AP preregistration is not explained yet; citation?--> scheme that builds upon a reproducible workflow, the so-called _preregistration as code_ (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript as dynamic document, including the sections Introduction, Methods, and Results before they gather empirical data.
This dynamic document closely resembles an actual journal article but with "fake" results from simulated data that serve to generate placeholders for figures, tables, and statistics. 
This document serves as unambiguous preregistration of the planned analysis and how the results are going to be reported.
The only parts that are missing from this first draft are the discussion of results, which will completed once the real data were gathered.

@@

We argue that the benefits of open science increase if research is not only transparent but reproducible.
To this end, this tutorial demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
A rigorous standard increases the likelihood of any individual project to be actually reproducible and remain reproducible as long as possible.
An easy-to-use standard, on the other hand, is less error-prone, produces less overhead and is thus more likely to be adopted.
To us, it makes sense to strive for a broad adoption of easy-to-use practices and then gradually move into more rigorous standards as researchers become more proficient with the tools involved.
The remainder of this paper is structured along these lines.
The first part introduces theoretical concepts and software solutions that together form a rigorous reproducibility standard.
The second part is a tutorial that shows how these software solutions can be easily employed by researchers with the help of the R package `repro`.

We have structured the tutorial with a _learning by doing_ approach in mind, such that readers can follow along on their own computers.
We explicitly encourage readers to try out all R commands for themselves.
Unless stated otherwise, all code blocks are meant to be run in the statistical programming language R (tested with version `r with(R.version, paste0(major, ".", minor))`).

@@

1.	**Multiple inconsistent versions of code, data, or both**; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or the analysis code could have been modified during writing of a paper because a bug was removed at some point in time; it may then be unclear, which version of code and data was used to produce some reported set of results.
2.	**Copy-and-paste errors**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., copy-pasting results from one program to another; reverse coding items on the fly)
4.	**External code dependencies**; for example, a given analysis may depend on a specific version of a specific software package that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
<!-- CJ 5. Missing steps: Some of the steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., copy-pasting results from one program to another; reverse coding items on the fly).-->
<!--AB Caspar's 5th point is important. In my mind, this best fits to point #3, which adressed an insufficient level of coding/automation/documentation of the necessary steps taken. Can we merge this into one new point #3? I will commit a suggested change of point #3 here (see above);.-->

@@

```{r}
repro::check_git()
repro::check_make()
repro::check_docker()
```

@@

Whenever you made changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
The easiest way to do this is to use the RStudio Git pane.
Click on the empty box next to the file you want to stage, and when a tick appears, which means that the file is staged.
After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.

@@

Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow the suggestions shown on the command line.

@@

The text of an R Markdown is written in Markdown. <!--AB citation needed for further information-->
The following Markdown example serves to illustrate the Markdown syntax. 
It shows how to create a heading, a word in bold font, and a list of several items in Markdown:

@@

Users can have full control over the software installed within the image of the project.
`repro` creates three Dockerfiles inside the `.repro` directory.
The file `.repro/Dockerfile_base` contains information about the base image on which all the remaining software is installed.
By default we rely on the "verse" images provided by the Rocker project [@boettigerIntroductionRockerDocker2017].
These contain (among other things) the packages `tidyverse`, `rmarkdown` and a complete LaTeX installation, which makes these images ideal for the creation of scientific manuscripts.
Users can choose which R version they want to have inside the container by changing the version number in line 1 to the desired R version number.
By default the R version corresponds to the locally installed <!--AB: please check - did I get this correctly? It is the local version?-->version on which `repro::automate()` was called the first time.
The build date is used to install packages in the version which was available on CRAN on this specific date and can also be changed, by default this date is set to the date on which `repro::automate()` was called the first time. 
This way, the call to the automate function virtually freezes the R environment to the state it was called the first time inside the container. 
Below, you see the Docker base file we used to create this manuscript:

@@

The last Dockerfile, `.repro/Dockerfile_packages`, is automatically generated from the metadata provided and must not be manually changed.
Whenever `repro::automate()` is called, repro gathers all R packages from all `.Rmd` files and figures out whether they should be installed from CRAN or GitHub. <!-- does this also figure out the specific version? If so, please say so-->

@@

The file `.repro/Makefile_Rmds` contains the automatically generated targets from `repro::automate()` for the R Markdown files.
This file should not be altered manually.
If you want to adept<!--AB strange word?! change--> a target than simply provide an alternative target in the main `Makefile`.
Targets in the main `Makefile` take precedent.

>>>>

This tutorial is aimed at researchers who intend to ensure the reproducibility of a research project and are willing to make relevant code, data, and materials available, whether publicly or on request.
We pursue two aims.
First, the present paper introduces the reader to a workflow that ensures reproducibility by building on several technical solutions originating in the software engineering community that eliminate four common threats to reproducibility [see @Peikert2019 for details].
Specifically, we demonstrate how the R package `repro` supports researchers in implementing these solutions.
The second objective is to discuss opportunities of such workflows to improve not only how research outputs are created and recreated but also how research itself is conducted.
One such opportunity is an alternative preregistration [@NosekRevolution2018] scheme that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript as dynamic document, including the sections Introduction, Methods, and Results before they gather empirical data.
This dynamic document closely resembles an actual journal article but with "fake" results from simulated data that serve to generate placeholders for figures, tables, and statistics. 
This document serves as unambiguous preregistration of the planned analysis and how the results are going to be reported.
The only parts that are missing from this first draft are the discussion of results, which the authors add after they gathered empirical data.

@@

We argue that the benefits of open science increase if research is not only transparent but reproducible.
To this end, this tutorial demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
A rigorous standard increases the likelihood of any individual project to be actually reproducible and remain reproducible as long as possible.
An easy-to-use standard, on the other hand, is less error-prone, produces less overhead and is thus more likely to be adopted.
To us, it makes sense to strive for a broad adoption of easy-to-use practices and then gradually move into more complex workflows as researchers become more proficient with the tools involved.
The remainder of this paper is structured along these lines.
The first part introduces theoretical concepts and software solutions that together form a rigorous reproducibility standard.
The second part is a tutorial that shows how these software solutions can be easily employed by researchers with the help of the R package `repro`.

We have structured the tutorial with a *learning by doing* approach in mind, such that readers can follow along on their own computers.
We explicitly encourage readers to try out all R commands for themselves.
Unless stated otherwise, all code blocks are meant to be run in the statistical programming language R (tested with version `r with(R.version, paste0(major, ".", minor))`).

@@

1.	**Multiple inconsistent versions of code, data, or both**; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or the analysis code could have been modified during writing of a paper because a bug was removed at some point in time; it may then be unclear, which version of code and data was used to produce some reported set of results.
2.	**Copy-and-paste errors**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., running a scripts/commands by hand; copy-pasting results from one program to another)
4.	**External code dependencies**; for example, a given analysis may depend on a specific version of a specific software package that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
<!-- CJ 5. Missing steps: Some of the steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., copy-pasting results from one program to another; reverse coding items on the fly).-->
<!--AB Caspar's 5th point is important. In my mind, this best fits to point #3, which adressed an insufficient level of coding/automation/documentation of the necessary steps taken. Can we merge this into one new point #3? I will commit a suggested change of point #3 here (see above);.-->

@@

```{r}
# `package::function()` → use function from package without `library(package)`
repro::check_git()
repro::check_make()
repro::check_docker()
```

@@

For Git to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
The easiest way to do this is to use the RStudio Git pane.
Click on the empty box next to the file you want to stage, and when a tick appears, which means that the file is staged.
After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.

@@

Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow the suggestions shown on the R console.

@@

The text of an R Markdown is written in Markdown. <!--AB citation needed for further information-->
The following example serves to illustrate the Markdown syntax. 
It shows how to create a heading, a word in bold font, and a list of several items in Markdown:

@@

Users can have full control over the software installed within the image of the project.
`repro` creates three Dockerfiles inside the `.repro` directory.
The file `.repro/Dockerfile_base` contains information about the base image on which all the remaining software is installed.
By default we rely on the "verse" images provided by the Rocker project [@boettigerIntroductionRockerDocker2017].
These contain (among other software) the packages `tidyverse`, `rmarkdown` and a complete LaTeX installation, which makes these images ideal for the creation of scientific manuscripts.
Users can choose which R version they want to have inside the container by changing the version number in line 1 to the desired R version number.
By default the R version corresponds to the locally installed version on which `repro::automate()` was called the first time.
The build date is used to install packages in the version which was available on CRAN on this specific date and can also be changed, by default this date is set to the date on which `repro::automate()` was called the first time. 
This way, the call to the automate function virtually freezes the R environment to the state it was called the first time inside the container. 
Below, you see the Docker base file we used to create this manuscript:

@@

The last Dockerfile, `.repro/Dockerfile_packages`, is automatically generated from the metadata provided and must not be manually changed.
Whenever `repro::automate()` is called, repro gathers all R packages from all `.Rmd` files and figures out whether they should be installed from CRAN or GitHub fixed to the date specified in `Dockerfile_base`.

@@

The file `.repro/Makefile_Rmds` contains the automatically generated targets from `repro::automate()` for the R Markdown files.
This file should not be altered manually.
If you are not satisfied with the automatically generated target than simply provide an alternative target in the main `Makefile`.
Targets in the main `Makefile` take precedent.
